---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Surveyexplorer

Visualize and tabulate single-choice, multiple-choice, matrix-style questions from survey data. Includes ability to group cross-tabulations, frequency distributions, and plots by categorical variables and to integrate survey weights. Ideal for quickly uncovering descriptive patterns in survey data.

## Installation

```{r eval=FALSE}
# install.packages("devtools"), if not already downloaded
devtools::install_github("liamhaller/surveyexplorer")
```

# Examples

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(surveyexplorer)
```

The data used in the following examples is from the `berlinbears` dataset, a fictional survey of bears in Berlin, that is included in the `surveyexplorer` package.

### Single-choice questions

```{r, eval= FALSE}
#Basic table
single_table(berlinbears, 
             question = income)
```

```{r, echo=FALSE}
single_table(berlinbears, 
             question = income) |> gt::as_raw_html()
```

Use `group_by =` to partition the question into several groups

```{r}
single_table(berlinbears,
             question = income,
             group_by = gender)
```

Ignore unwanted subgroups with `subgroups_to_exclude`

```{r}
single_table(berlinbears,
             question = income, 
             group_by = gender, 
             subgroups_to_exclude = NA) |> gt::as_raw_html()
```

Remove NAs from the question variable with `na.rm`

```{r}
single_table(berlinbears,
             question = income, 
             group_by = gender, 
             subgroups_to_exclude = NA,
             na.rm = TRUE)
```

Finally, you can specify survey weights using the weight option

```{r}
single_table(berlinbears,
             question = income, 
             group_by = gender, 
             subgroups_to_exclude = NA,
             na.rm = TRUE,
             weights = weights)
```

The same syntax applies to the `single_freq` function to plot frequencies of the question optionally partitioned by subgroups.

```{r, fig.width=10}
single_freq(berlinbears,
             question = income, 
             group_by = gender, 
             subgroups_to_exclude = NA,
             na.rm = TRUE,
             weights = weights)
```

### Multiple-choice questions

The options and syntax for multiple-choice tables `multi_table` and graphs `multi_graphs` are the same. The only difference is the question input also accommodates tidyselect syntax to sekect several columns for each answer option. For example, the question "will_eat" has five answer options each prefixed by "will_eat"

```{r}
berlinbears |> 
  dplyr::select(starts_with('will_eat')) |> 
  head()
```

The same syntax can be used to select the question for the multiple choice tables and graphs

```{r, results='hold'}
multi_table(berlinbears, 
            question = dplyr::starts_with('will_eat'), 
            group_by = genus, 
            subgroups_to_exclude = NA,
            na.rm = TRUE)
```

For graphing, the `multi_freq` function creates an UpSet plot to visualize the frequencies of the intersecting sets for each answer combination and also includes the ability to specify weights.

```{r}
multi_freq(berlinbears, 
            question = dplyr::starts_with('will_eat'), 
            na.rm = TRUE,
            weights = weights)
```

The graphs can also be grouped

```{r}
multi_freq(berlinbears, 
            question = dplyr::starts_with('will_eat'), 
            group_by = genus,
            subgroups_to_exclude = NA,
            na.rm = FALSE,
            weights = weights)
```

### Matrix Questions

`matrix_table` has the same syntax as above and works with array or categorical questions

```{r}
matrix_table(berlinbears, 
             dplyr::starts_with('p_'),
             group_by = is_parent)
matrix_table(berlinbears, 
             dplyr::starts_with('c_'),
             group_by = is_parent)
```

`matrix_freq` visualizes the frequencies of responses

```{r}
matrix_freq(berlinbears, 
             dplyr::starts_with('p_'))
matrix_freq(berlinbears, 
             dplyr::starts_with('c_'),
             group_by = is_parent)
```

For array/matrix style questions that are numeric `matrix_mean` plots the mean values and confidence intervals

```{r}
matrix_mean(berlinbears, 
             question = dplyr::starts_with('p_'),
             na.rm = TRUE)
#Can also apply grouping
matrix_mean(berlinbears, 
            question = dplyr::starts_with('p_'),
            na.rm = TRUE,
            group_by = species, 
            subgroups_to_exclude = NA)
#with survey weights
matrix_mean(berlinbears, 
            question = dplyr::starts_with('p_'),
            na.rm = TRUE,
            group_by = species, 
            subgroups_to_exclude = NA, 
            weights = weights)
```

Finally, for Likert questions (scales of 3,5,7,9...) `matrix_likert` provides a custom plot

```{r}
matrix_likert(berlinbears, 
              question = dplyr::starts_with('p_'))
#you can specify custom labels with the `label` argument
matrix_likert(berlinbears,
              question = dplyr::starts_with('p_'),
              labels = c('Strongly disagree', 'Disagree','Neutral','Agree','Strongly agree'))

#and pass colors using the colors option 
matrix_likert(berlinbears, 
              question = dplyr::starts_with('p_'),
              labels = c('Strongly disagree', 'Disagree','Neutral','Agree','Strongly agree'),
              colors = c("#E1AA28", "#1E5F46", "#7E8F75", "#EFCD83", "#E17832"))
#can also apply weights 
matrix_likert(berlinbears, 
              question = dplyr::starts_with('p_'),
              labels = c('Strongly disagree', 'Disagree','Neutral','Agree','Strongly agree'), 
              colors = c("#E1AA28", "#1E5F46", "#7E8F75", "#EFCD83", "#E17832"),
              weights = weights) 
```

## Overview

### Functions

-   Single-choice
    -   `single_table`
    -   `single_freq`
-   Multiple-choice
    -   `multi_table`
    -   `multi_freq`
-   Matrix
    -   `matrix_table`
    -   `matrix_freq`
    -   `matrix_mean`
    -   `matrix_likert`

\*\_table functions return a gt table of the cross tabulations and frequencies for each question while \*\_freq returns the same data but as a plot.

For matrix-style questions with numerical input, `matrix_mean` plots the mean value value and ± two standard deviations. `matrix_likert` visualizes questions that accept Likert responses (strongly agree-strongly disagree) or questions with 3,5,7,9... categories.

### Syntax

Each function contains the following options

-   dataset ---The input dataframe (or tibble) of survey questions
-   question --- The column(s) that contain the response options for a question, can be selected by using tidyselect semanatics or providing a vector of column names or numbers
-   group_by --- Optional variable to group the analysis. If provided, the frequencies and counts will be calculated within each subgroup
-   subgroups_to_exclude --- Optional vector specifying subgroups to exclude from the analysis
-   weights --- Optional variable containing survey weights. If provided, frequencies and counts will be weighted accordingly
-   na.rm --- Logical indicating whether to remove NA values from question before analysis
